Cooperative Control for Narrow Passage Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station

### 2012-10-10 manuscript.tex

%
% ICRA 2013 Paper Submission
% created on Sep 5, 2012
%
\def \papertitle{Cooperative Control for Window Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}

% NOTE: WE WRITE ALL TEXT IN FIRST PERSON PLURAL ("we")
% Conventions: flapping-wing, small-scale

%!TEX TS-program = latex
%!TEX encoding = UTF-8 Unicode

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts % for \thanks
\overrideIEEEmargins % for \addtolength

% Packages
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[pdftex]{hyperref}

% hyperref setup
\definecolor{linkCol}{gray}{0.25}
\hypersetup{
    pdftitle={\papertitle},
    pdfauthor={Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing},
    pdfsubject={},
    pdfkeywords={},
    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
}

% Title
\title{\LARGE \bf \papertitle}

% Authors
\author{
    Ryan C. Julian$^{*}$,
    Cameron J. Rose$^{*}$,
    Humphrey Hu$^{\dagger}$, and
    Ronald S. Fearing$^{*}$
    \thanks{$^{*}$
    	R.C. Julian, C.J. Rose and R.S. Fearing are
        with the Department of Electrical Engineering and Computer Sciences,
        University of California, Berkeley, CA 94720 USA.
        \href{mailto:ryanjulian@berkeley.edu}{\tt ryanjulian@berkeley.edu},
        \href{mailto:c\_rose@eecs.berkeley.edu}{\tt c\_rose@eecs.berkeley.edu},
        \href{mailto:ronf@eecs.berkeley.edu}{\tt ronf@eecs.berkeley.edu}
    }
    \thanks{$^{\dagger}$
        H. Hu is with the Robotics Institute,
        Carnegie Mellon University, Pittsburgh, PA 15213 USA.
        \href{mailto:humphreh@andrew.cmu.edu}{\tt humphreh@andrew.cmu.edu}
    }
    \thanks{
        This material is based upon work supported by the National Science
        Foundation under Grants No. CNS-0931463, and by the Army Research
        Laboratory under the Micro Autonomous System and Technology
        Collaborative Technology Alliance.
    }
}

@@

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%TODO(ryan)
The power, size, and weight constraints of micro air vehicle (MAV) platforms limit their on-board sensing and computational resources. In this work, we demonstrate cooperative control of a 14 gram ornithopter MAV by a lightweight ground station using computer vision. We also introduce a new ornithopter MAV platform, the h$^2$Bird, which...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%TODO(ryan)
The power, size, and weight constraints of micro air vehicle (MAV) platforms significantly limit on-board sensing and computational resources, and thus the complexity of the control and state estimation algorithms which can be executed on-board. These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
%TODO(humphrey)

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.jpg}
\caption{H$^2$Bird Ornithopter MAV}
\label{fig:h2bird}
\end{figure}

The flying platform used in this experiment is the H$^2$Bird flapping wing MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams fully loaded. A tail propeller and servo-controlled tail elevator provide control in the yaw and pitch axes. An onboard electronics package combines a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor drivers with a 90 mAh lithium polymer battery for autonomous flight. In routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and can operate for approximately [10] minutes.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird_axes.jpg}
\caption{H$^2$Bird Ornithopter with flight control axes}
\label{fig:h2Bird_axes}
\end{figure}

@@

\begin{equation}
\label{quat_error}
q_e = q'\otimes q_r
\end{equation}

\begin{equation}
\label{quat_angle}
\alpha = 2acos(q_{e,w})
\end{equation}

\begin{equation}
\label{quat_linearize}
Q_e = \alpha*q_e/sin(\alpha /2)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
%TODO(ryan)

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/window_trials.jpg}
\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
\label{fig:experiment_cone}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addtolength{\textheight}{-0cm}
                        % This command serves to balance the column lengths
                        % on the last page of the document manually. It shortens
                        % the textheight of the last page by a suitable amount.
                        % This command does not take effect until the next page
                        % so it should come on the page before the last. Make
                        % sure that you do not shorten the textheight too much.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
The authors would like to thank Fernando Garcia Bermudez for his assistance with the Vicon motion capture system, Andrew Pullin for his help with robot photography, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtranS}
\bibliography{}

\end{document}

>>>>

% This is "aamas2013 .tex" September 2012 
% This file should be compiled with "aamas2013 .cls" 
% This example file demonstrates the use of the 'aamas2013 .cls'
% LaTeX2e document class file. It is for those submitting
% articles to AAMAS 2013  conference. This file is based on
% the sig-alternate.tex example file.
% The 'sig-alternate.cls' file of ACM will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
% than the original style ACM style.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls ) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with AAMAS data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'aamas2013 .cls' you don't have control
% from within the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the IFAAMAS Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% These information will be overwritten by fixed AAMAS 2013  information
% in the style files - it is NOT as you are used with ACM style files.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%

% This is the document class for full camera ready papers and extended abstracts repsectively

\documentclass{aamas2013}

% if you are using PDF LaTex and you cannot find a way for producing
% letter, the following explicit settings may help

\pdfpagewidth=8.5truein
\pdfpageheight=11truein

@@

% In the original styles from ACM, you would have needed to
% add meta-info here. This is not necessary for AAMAS 2013  as
% the complete copyright information is generated by the cls-files.

\title{Cooperative Control for Narrow Passage Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}

% AUTHORS

% For initial submission, do not give author names, but the
% tracking number, instead, as the review process is blind.

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%

\numberofauthors{4}

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
% Paper number
%\alignauthor Paper  XXX
% 1st. author
\alignauthor Ryan C. Julian\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{ryanjulian@berkeley.edu}
% 2nd. author
\alignauthor Cameron J. Rose\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{c\_rose@eecs.berkeley.edu}
% 3rd. author
\alignauthor Humphrey Hu\\
       \affaddr{Robotics Institute}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA 15213}\\
       \email{humhu@cmu.edu}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Ronald S. Fearing\\
       \affaddr{Department of EECS}\\
       \affaddr{Univ. of California, Berkeley}\\
       \affaddr{Berkeley, CA 94720}\\
       \email{ronf@eecs.berkeley.edu}
}

%% There's nothing stopping you putting the seventh, eighth, etc.
%% author on the opening page (as the 'third row') but we ask,
%% for aesthetic reasons that you place these 'additional authors'
%% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
%% Just remember to make sure that the TOTAL number of authors
%% is the number that will appear on the first page PLUS the
%% number that will appear in the \additionalauthors section.

\maketitle

\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit their on-board sensing and computational resources. Ground vehicles have less mobility than MAVs, but relaxed size constraints and typically more computing power, creating opportunities for cooperation. In this work, we demonstrate cooperative target-seeking between a 13 gram ornithopter MAV and a lightweight ground station using computer vision.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It features clap and fling wings, improves upon a previous design by utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a payload of 2.8 grams. We augment the ornithopter's built-in gyroscope-based control with a ground station, which has power and weight requirements appropriate for deployment on a ground vehicle. The ground station provides heading estimates to the ornithoper by running a real-time motion tracking algorithm on a live video stream. This allows the ornithopter to reliably negotiate flight through narrow openings.

We demonstrate the performance of the location estimation and navigation algorithms. We model the backwards-reachable region for successfully navigating through narrow openings. Finally, we verify this model through with collected in experiments with window traversal.
\end{abstract}

% Note that the category section should be completed after reference to the ACM Computing Classification Scheme available at
% http://www.acm.org/about/class/1998/.

\category{I.2.9}{Robotics}{Autonomous Vehicles}
\category{I.2.9}{Robotics}{Distributed Artificial Intelligence}[Multiagent systems] 
\category{I.2.9}{Robotics}{Vision and Scene Understanding}[Motion]

%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%General terms should be selected from the following 16 terms: Algorithms, Management, Measurement, Documentation, Performance, Design, Economics, Reliability, Experimentation, Security, Human Factors, Standardization, Languages, Theory, Legal Aspects, Verification.

\terms{Algorithms, Performance, Design, Experimentation}

%Keywords are your own choice of terms you would like the paper to be indexed by.

\keywords{AAMAS proceedings, multiagent systems, cooperative control, particle filter, micro air vehicle, ground station, visual servoing, pose estimation, flapping wing, ornithopter, biomimetic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%TODO(ryan)
The power, size, and weight constraints of micro air vehicle (MAV) platforms significantly limit on-board sensing and computational resources, and thus the complexity of the control and state estimation algorithms which can be executed on-board. These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird Ornithopter MAV}
\label{fig:h2bird}
\end{figure}

@@

%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper MAV}
%TODO(humphrey)
The flying platform used in this experiment is the H$^2$Bird flapping wing MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams fully loaded. A tail propeller and servo-controlled tail elevator provide control in the yaw and pitch axes. An onboard electronics package combines a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor drivers with a 90 mAh lithium polymer battery for autonomous flight. In routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and can operate for approximately [10] minutes.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=\linewidth]{figures/h2bird_axes.png}
\includegraphics[height=150pt]{figures/h2bird_axes.png}
\caption{H$^2$Bird Ornithopter with flight control axes}
\label{fig:h2Bird_axes}
\end{figure}

@@

\begin{equation}
\label{quat_error}
q_e = q'\otimes q_r
\end{equation}
\begin{equation}
\label{quat_angle}
\alpha = 2acos(q_{e,w})
\end{equation}
\begin{equation}
\label{quat_linearize}
Q_e = \alpha*q_e/sin(\alpha /2)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance}
%TODO(cameron)

@@

%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
%TODO(cameron)

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/flight_paths.pdf}
\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
\label{fig:flight_paths}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
The authors would like to thank Fernando Garcia Bermudez for his assistance with the Vicon motion capture system, Andrew Pullin for his help with robot photography, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% The following two commands are all you need in the
%% initial runs of your .tex file to
%% produce the bibliography for the citations in your paper.
%\bibliographystyle{abbrv}
%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
%% You must have a proper ".bib" file
%%  and remember to run:
%% latex bibtex latex latex
%% to resolve all references
%%
%% ACM needs 'a single self-contained file'!
%%
%%APPENDICES are optional
%%\balancecolumns
%\appendix
%%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
%% This next section command marks the start of
%% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The aamas2013 .cls file is based on the
%sig-alternate.cls file that is itself chock-full of succinct and
%helpful comments.  If you consider yourself a moderately experienced
%to expert user of \LaTeX, you may find reading it useful but please
%remember not to change it.
%%\balancecolumns % GM June 2007
%% That's all folks!
\end{document}
